Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-07-05 PandaEvolution.tex
Address Shantenu's comments

Supercomputers are designed to support parallel computation that
requires runtime communication. Jobs are executed across multiple nodes, each
node calculating a small part of the problem and communicating with other
node, typically via MPI\@. Accordingly, supercomputers have large number of worker
nodes, connected through a high-speed, low-latency dedicated network. Each
worker node has multicore CPUs, usually augmented with Graphics Processing
Units (GPUs) or other specialized co-processors.

@@

% The Titan supercomputer, current number three on the Top 500
% list,%~\cite{top500}, 
% is a Cray XK7 system with 18,688 worker nodes and a total
% of 299,008 AMD cores and 18,688 Nvidia Tesla GPUs. 
% Each worker node has an AMD Opteron 6274 16-core CPU,
% a Nvidia Tesla K20X GPU, 32 GB of RAM and no local storage, though a 16 GB
% RAM disk can be set up. Work nodes use Cray’s Gemini interconnect for
% inter-node MPI messaging. Titan is served by the Spider
% II~\cite{oral2013olcf}, a Lustre filesystem with 32 PB of disk storage, and
% by a 29 PB HPSS tape storage system. Titan’s worker nodes run Compute Node
% Linux, a run time environment based on SUSE Linux Enterprise Server.

% Titan's users submit jobs to Titan's PBS scheduler by logging into login or
% data transfer nodes (DTNs). Titan's authentication and authorization model
% is based on two-factor authentication with a RSA SecurID key. Login nodes
% and DTNs have out/inbound wide area network connectivity while worker nodes
% have only local network access. Fair-share and allocation policies are in
% place both for the PBS batch system and shared file systems.

@@

Currently, very few HEP applications can benefit from Titan's GPUs but some
computationally-intensive and non memory-intensive tasks of ATLAS workflows
can be off-loaded from the Grid to Titan's. Further, when HEP tasks can be
partitioned into independent jobs, Titan worker nodes can be used to execute
up to 16 concurrent payloads, one per each available core. Given these
constraints and challenges, the  Monte Carlo detector simulation task is most
suitable for execution on Titan at the moment. This type of task is mostly
computational-intensive, requiring less than 2GB of RAM at runtime and small
input data. Detector simulation tasks in ATLAS are performed via
AthenaMP~\cite{aad2010atlas}, the ATLAS software framework integrating the
GEANT4 detector simulation toolkit~\cite{agostinelli2003geant4}. These tasks
account for $\approx$ 60\% of all the jobs on WLCG, making them a primary
candidate for offloading.

>>>>

% Supercomputers are designed to support parallel computation that requires
% runtime communication. Jobs are executed across multiple nodes, each node
% calculating a small part of the problem and communicating with other node,
% typically via MPI\@. Accordingly, supercomputers have large number of
% worker nodes, connected through a high-speed, low-latency dedicated
% network. Each worker node has multicore CPUs, usually augmented with
% Graphics Processing Units (GPUs) or other specialized co-processors.

@@

% The Titan supercomputer, current number three on the Top 500
% list,%~\cite{top500}, 
% is a Cray XK7 system with 18,688 worker nodes and a total
% of 299,008 AMD cores and 18,688 Nvidia Tesla GPUs. 
% Each worker node has an AMD Opteron 6274 16-core CPU,
% a Nvidia Tesla K20X GPU, 32 GB of RAM and no local storage, though a 16 GB
% RAM disk can be set up. Work nodes use Cray’s Gemini interconnect for
% inter-node MPI messaging. Titan is served by the Spider
% II~\cite{oral2013olcf}, a Lustre filesystem with 32 PB of disk storage, and
% by a 29 PB HPSS tape storage system. Titan’s worker nodes run Compute Node
% Linux, a run time environment based on SUSE Linux Enterprise Server.
%
% Titan's users submit jobs to Titan's PBS scheduler by logging into login or
% data transfer nodes (DTNs). Titan's authentication and authorization model
% is based on two-factor authentication with a RSA SecurID key. Login nodes
% and DTNs have out/inbound wide area network connectivity while worker nodes
% have only local network access. Fair-share and allocation policies are in
% place both for the PBS batch system and shared file systems.

@@

Currently, very few HEP applications can benefit from Titan's GPUs but some
computationally-intensive and non memory-intensive tasks of ATLAS workflows
can be off-loaded from the Grid to Titan. Further, when HEP tasks can be
partitioned into independent jobs, Titan worker nodes can be used to execute
up to 16 concurrent payloads, one per each available core. Given these
constraints and challenges, the  Monte Carlo detector simulation task is most
suitable for execution on Titan at the moment. This type of task is mostly
computational-intensive, requiring less than 2GB of RAM at runtime and small
input data. Detector simulation tasks in ATLAS are performed via
AthenaMP~\cite{aad2010atlas}, the ATLAS software framework integrating the
GEANT4 detector simulation toolkit~\cite{agostinelli2003geant4}. These tasks
account for $\approx$ 60\% of all the jobs on WLCG, making them a primary
candidate for offloading.
